---
title: "assignment1"
author: "Jack Carbaugh"
date: "9/22/2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
This loads in the library Tidyverse
```{r}
library(tidyverse)
```

Create the mtcars dataframe
```{r}
# call built-in data mtcars.
data(mtcars)
```

filter out the row entries where mpg < 20
```{r}
# Select only car models where mpg<20
mtcars_mpg2 <- mtcars[mtcars$mpg < 20,]
```

Limit columns to just mpg, cyl, disp, hp, gears
```{r}
# Reduce the variables to mpg, cyl, disp, hp, gears
mtcars_mpg2 <- mtcars_mpg2[, c(1,2,3,4,10)]
```

Source the hand_functions.R file so that we can utilize its functions
```{r}
# read the R file hand_functions.R so that it can be used
# notice that with echo = TRUE 
source(file = "hand_functions.R", echo = TRUE)
```

Uses the sum_special function that is present in hand_functions.R. This function takes in a dataframe is input. If the input is not a dataframe, the function is halted and an error meassge is given. Otherwise, the function gets the mean, variance, covariance, and correlation of the dataframe, adds each statistic into a list, and returns said list. We can also tun the debugger line by line (while using line breaks) to test our code.

```{r}
# Now use the function from hand_functions.R
sp_out <- sum_special(mtcars_mpg2)
sp_out
```

Load esquisse, allowing us to build a ggplot for mtcars with drag n drop
```{r}
# library(esquisse)
# 
# esquisser(data = mtcars_mpg2, viewer = "browser")
```

Biulds a smooth ggplot showing the relationship between displacement and mpg.
```{r}
ggplot(mtcars_mpg2) +
  aes(x = disp, y = mpg) +
  geom_point(shape = "bullet", size = 4L, colour = "#B22222") +
  geom_smooth(span = 0.5) +
  theme_minimal()
```

Boxplot showing mpg, factored out by the cylinder type
```{r}
# note that this boxplot cannot be made with esquisse() unless
# the data is adjusted.  What adjustment is needed?


ggplot(mtcars_mpg2, aes(x=as.factor(cyl), y=mpg)) + 
  geom_boxplot(fill="slateblue", alpha=0.2) + 
  xlab("cyl")
```

